---
title: "VIBER Vignette"
author: "Giulio Caravagna"
date: "14 February 2019"
institute: "Institute for Cancer Research"
email: "giulio.caravagna@ed.ac.uk"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteEngine{knitr::knitr}
  %\VignetteIndexEntry{VIBER vignette}
  %\usepackage[UTF-8]{inputenc}
---
```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



```{r, message=FALSE, warning=FALSE}
library(VIBER)
library(tidyverse)
```

To run the fit you need 2 tibbles (or matrices), with equal dimensions and matched column names. One tibble contains the *counts of the successful Bernoulli trials*, the other contains the *count of all the attempted trials*. 

In the package we provide `mvbmm_example`, a dataset which shows the input format.

```{r}
data("mvbmm_example")
```
The format is (`S1` and `S2` are the *dimensions*)
```{r}
mvbmm_example$successes
```

## Fitting the model

You can fit the model with the `variational_fit` function, which is quite well documented
```{r}
??variational_fit
```

You have control over several parameters. Concerning the mixture type you can set the following 

- `K`, the maximum number of clusters returned
- `alpha_0`, the concentration parameter of the Dirichlet mixture. 
- `a_0` and `b_0`, the prior Beta hyperparameter for each Binomial component

Concerning the variational optimization you can set the following

- `max_iter`, the maximum number of fit iterations
- `epsilon_conv`, the epsilon to measure convergence as ELBO absolute difference
- `samples`, the number of fits computed. 

The fitting engine makes use of the [easypar](https://github.com/caravagn/easypar) package to run in parallel the required number of fits. The default is multi-core implementation that uses 80% of the available cores; you can disable parallel runs and execute sequential turning off `easypar`, as explained in its [Wiki](https://github.com/caravagn/easypar).

We run the fit with default parameters, the output model is print to screen.
```{r}
fit = variational_fit(
  mvbmm_example$successes,
  mvbmm_example$trials
)
```

## Filtering output clusters

Because the model is semi-parametric, it will attempt to use at most `K` Binomial clusters. However, most of those clusters might be not interesting and you might want to filter them.

VIBER implements 2 possible filters which are available in function `choose_clusters`.


- Filter 1 [F1] uses the cluster size (mixing proportion), which one can be required to be above a certain cutoff;
- Filter 2 [F2] the Binomial peaks, which one can ask to be above a certain value (e.g., $p>0.05$) in at least a certain number of dimensions.

Adter filtering, output clusters will be renamed by size (C1 will be larger etc.), and the latent variables and  hard clustering assignments will be updated accordingly.

Here we require only the minimum cluster size to be 2% of the total number of points.
```{r}
fit = choose_clusters(fit, 
                      binomial_cutoff = 0, 
                      dimensions_cutoff = 0,
                      pi_cutoff = 0.02)
```
The new model has fewer clusters
```{r}
fit 
```


## Plots

You can plot the data - one dimension against the other - with the `plot_2D` (for instance, trye `plot_2D(fit, d1 = 'S1', d2 = 'S2')`), or use the S3 function `plot(fit)` to compute a list of plots for each pair of dimensions in the mixture.

```{r, fig.height=3, fig.width=3}
plot(fit)
```


You can plot the ELBO
```{r, fig.height=3, fig.width=3}
plot_ELBO(fit)
```

You can plot the Binomial peaks, per cluster and per dimension
```{r, fig.height=3, fig.width=5}
plot_peaks(fit)
```

You can plot the mixing proportions of the mixture 
```{r, fig.height=3, fig.width=3}
plot_mixing_proportions(fit)
```

And, finally, you can plot the latent variables of the mixture 
```{r, fig.height=5, fig.width=4}
plot_latent_variables(fit)
```

